---
title: "HW2_Base_R"
author: "Chris Hill"
format: html
editor: visual
---

## Task 1: Basic Vector Practice

Create two vectors. One vector corresponding to the pre measurements and one to the post measurements.

Assign names to the vector elements using the paste() function.

```{r}
pre_treatment <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)

post_treatment <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)

patient_names <- paste("Subject", 1:20, sep = "_")

names(pre_treatment) <- patient_names
names(post_treatment) <- patient_names
```

```{r}
pre_treatment
```

```{r}
post_treatment
```

Calculate the change in blood pressure for each patient by subtracting post-treatment measurements from pre-treatment measurements.

```{r}
bp_change <- pre_treatment - post_treatment
bp_change
```

Calculate the average decrease in blood pressure across all patients

```{r}
mean_change <- mean(bp_change)
mean_change
```

Determine which patients experienced a decrease in blood pressure after treatment (a positive change). Use the which() function to just return the indices (and names) associated with this type of change.

```{r}
patients_with_decrease <- which(bp_change > 0)

patients_with_decrease_names <- names(bp_change)[patients_with_decrease]

patients_with_decrease_names
```

Subset the vector of differences to only return those that have a positive change. 

```{r}
bp_decreased <- bp_change[bp_change > 0]

bp_decreased
```

Calculate the average decrease in blood pressure for those where the blood pressure decreased (positive
change).

```{r}
mean_bp_decrease <- mean(bp_decreased)

mean_bp_decrease
```

## Task 2: Basic Data Frame Practice

Create a data frame object with four columns corresponding to your data above: 

- patient
- pre_bp
- post_bp
- diff_bp

```{r}
bp_df <- data.frame(
  patient = patient_names,
  pre_bp = pre_treatment,
  post_bp = post_treatment,
  diff_bp = pre_treatment - post_treatment
)

bp_df
```

Return only rows of the data frame where the diff_bp column is negative. (Use [ or learn about the subset() function if you’d like. If you use [, don’t reference the original vector from the first part, access the column of the data frame to make your comparison with 0.)

```{r}
negative_diff_rows <- bp_df[bp_df$diff_bp < 0, ]

negative_diff_rows
```

```{r}
negative_diff_rows_subset <- subset(bp_df, diff_bp < 0)

negative_diff_rows_subset
```

Add a new column to the data frame corresponding to TRUE if the post_bp is less than 120 and FALSE
if it is not. Similar to the previous question, don’t reference the original vector from task 1, access the
column of the data frame to make your comparison

```{r}
bp_df$below_120 <- bp_df$post_bp < 120

bp_df
```

Print the data frame out nicely in your final document

```{r}
knitr::kable(bp_df)
```

## Task 3: List Practice

Create a new data frame with the given placebo data in it that is similar to the data frame from task 2 (including the new column of TRUE/FALSE values).

```{r}
pre_treatment_placebo <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
post_treatment_placebo <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

subject_names_placebo <- paste("Subject", 1:10, sep = "_")

placebo_df <- data.frame(
  patient = subject_names_placebo,
  pre_bp = pre_treatment_placebo,
  post_bp = post_treatment_placebo,
  diff_bp = pre_treatment_placebo - post_treatment_placebo,
  below_120 = pre_treatment_placebo < 120
)

placebo_df
```

Now create and store a list with two elements:

- 1st element named treatment and contains the first data frame you created.
- 2nd element named placebo and contains the second data frame you created.

```{r}
experiment_list <- list(
  treatment = bp_df,
  placebo = placebo_df
)

str(experiment_list)
```

Access the first list element using different types of syntax (the result can be a list or just the data frame).

```{r}
experiment_list[[1]]
```

```{r}
experiment_list$treatment
```

```{r}
experiment_list[["treatment"]]
```

```{r}
experiment_list[1]
```

Use your list from question 2. In one line of code, access the pre_bp column of the placebo data frame.

```{r}
experiment_list$placebo$pre_bp
```

